---
title: "TT2021-05-25"
author: "Stéphanie Depickère (@sdepickere)"
date: "26/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidytuesdayR)
library(tidyverse)
library(ggpubr)
library(cowplot)
```

## TidyTuesday data: Mario Kart 64 World Records

The data this week comes from Mario Kart World Records (https://mkwrs.com/). 

Despite it’s release back in 1996 (1997 in Europe and North America), Mario Kart 64 World Records is still actively played by many and new world records are achieved every month.

The game consists of 16 individual tracks and world records can be achieved for the fastest single lap or the fastest completed race (three laps). Also, through the years, players discovered shortcuts in many of the tracks. 


## Download the data

```{r data, message=FALSE}
# load data
tt_data <- tt_load("2021-05-25")

# attribute the dataset to rr
rr <- tt_data$records
```

## Data preparing

The data contains a lot of details, with the records according to the tracks, the use of shortcuts, in single or 3 laps, etc. I was very interested in how the use of shortcuts could improve the records in the 3 laps according to the track.

```{r datazip, results='hide', warning=FALSE}
rr3 <- rr %>% filter(type=="Three Lap")
```

## plot

In this plot, we can see the evolution of the records obtained in the 16 tracks, with the use or not of shortcuts. 

```{r plot, fig.width=15, fig.height=12, warning=FALSE}
# choice of colors for the tracks
mycolors <- c("black","yellow","#26C4EC","magenta","#1E7FCB","orange","green","#BA9B61",
              "brown","violet","red","lightgreen","grey","pink","#0000FF","#B0F2B6")
# name for facets
shortcut.names <- as_labeller(
     c(`No` = "without shortcut", 
       `Yes` = "with shortcut"))

g <- ggplot(rr3,aes(date,time,color=track)) +
     geom_line(size=1.2) +
     facet_grid(cols=vars(shortcut),labeller = shortcut.names) +
     #     scale_color_viridis(discrete = TRUE, option = "D") +
     scale_color_manual(values = mycolors) +
     labs(x = NULL, y = "time (s)") +
     theme(strip.text.x = element_text(size=20, face="bold"),
           strip.background = element_rect(colour = "black", fill = "grey"),
           panel.background = element_rect(fill = "white", colour = "white"),
           panel.border = element_rect(fill = NA),
           panel.grid.major = element_line(colour = "grey",linetype="dotted"),
           axis.text.y = element_text(size=18,face="bold"),
           axis.text.x = element_text(size=18,face="bold"),
           axis.title.y = element_text(size=20,face="bold"),
           legend.title = element_text(size=20,face="bold"),
           legend.title.align = 0.5,
           legend.text = element_text(size=16)) 
g <- ggdraw(add_sub(g, "On the left, we can appreciate the increasing ability of the gamers who constantly improve the time needed to complete the race between 
                        \n1997 and 2021.On the right, we see how the discovering of shortcuts lowers dramatically the times. They have been discovered 
                        \nvery quickly, in 1997 for most of the tracks. For Choco Mountain or Rainbow Road tracks, different shortcuts were discovered over the 
                    \nyears. In the case of Rainbow Road, the record time was divided by 6!!",
               x = 0, hjust = 0,size=16,color="#303030",fontface="italic",lineheight = 0.5))




annotate_figure(g,
                top = text_grob("Mario Kart 64 World Records: improvement of records using shortcuts \n", 
                                color = "black", face = "bold", size = 25),
                bottom = text_grob("Data source: Mario Kart 64 World Records, Visualization: Stéphanie Depickère, May 2021", 
                                   color = "#2F1E0E",
                                   face = "italic", size = 15,hjust = 1, x = 1))

# save the plot
ggsave(filename = "MyTT2021-05-25.png",width = 15, height = 12,device = "png")

```
